A 38 Million Year Old Mini-Neptune in the Kepler Field

### 2021-09-22 paper/ms.tex
update ms

\keywords{
  exoplanet evolution (491),
  stellar associations (1582),
  open star clusters (1160),
	stellar ages (1581),
}

@@

The K2 and TESS missions have now enabled the detection of about ten
close-in planets younger than 100 million years, all smaller than
Jupiter
\citep{Mann_K2_33b_2016,David_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,plavchan_planet_2020,rizzuto_tess_2020,martioli_aumicbc_2021}.
The Kepler mission however has not yielded any planets with precise
ages below one gigayear \citep{Meibom_et_al_2013}.  The reason is that
during the main Kepler mission (2009--2013), only four open clusters
were known in the Kepler field: NGC\,6866, NGC\,6811, NGC\,6819, and
NGC\,6791, with ages spanning 0.7\,Gyr to 9\,Gyr
\citep{meibom_kepler_2011}.  Though isochronal and gyrochronal
analyses suggest that younger Kepler planets do exist
\citep{berger_identifying_2018,david_sizes_2021}, accurate and precise
age measurements typically require an ensemble of stars.  Fortunately,
recent analyses of the kinematic, photometric, and astrometric Gaia
data have significantly expanded our knowledge of cluster memberships
\citep[{\it
e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021,Kerr2021}.
As part of our Cluster Difference Imaging Photometric Survey (CDIPS,
\citealt{bouma_cdipsI_2019}), we concatenated the available analyses
from the literature, which yielded a list of candidate young and
age-dated stars (see Appendix~\ref{app:targetlist}).

@@

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Kepler and TESS views of the $\delta$\,Lyr cluster.} Colored
    points are kinematically selected candidate $\delta$\,Lyr cluster
    members.  Both Kepler (gray panels) and TESS observed portions of
    the cluster.  Naked-eye stars ($m_{\rm V}<6.5$) are shown in gray
    or orange.  Kepler\,1627 (green star) was observed during the
    entirety of the Kepler mission.  Colors of candidate cluster
    members indicate the duration of TESS observations to date, in
    units of lunar months.
    \label{fig:skychart}
	}
\end{figure}

@@

		\subfloat{
			\includegraphics[width=\textwidth]{f5.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The light curve of Kepler\,1627.}
    {\it Top}: 
    The full Kepler dataset spans 1{,}437 days (3.9 years), sampled at
    30 minute cadence.  A one hundred day segment is shown here.  The
    top panel shows the \texttt{PDCSAP} median-subtracted flux in
    units of parts-per-thousand ($\times 10^{-3}$).  The dominant
    signal is induced by starspots.  The model for the stellar
    variability (orange line) is subtracted below, revealing the
    transits of \pn, as well as other deviations from the model.  The
    online Figure Set shows the entire 3.9 years of observations.
    {\it Bottom}:
    Phase-folded transit of Kepler 1627Ab, with stellar variability
    removed.  The 2-$\sigma$ model uncertainties and the maximum {\it
    a posteriori} model are shown as the faint purple band, and the
    dark purple line.  Gray points are individual \texttt{PDCSAP} flux
    measurements; black points are binned to 20 minutes, with a
    representative error bar shown.  The asymmetric residual visible
    during the transit is robust against detrending methods; we
    believe that it is caused by starspot crossings.
    \label{fig:lc}
  }
\end{figure*}

@@

		\subfloat{
			\includegraphics[width=0.49\textwidth]{f16.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Evidence for a prograde orbit of Kepler 1627 Ab.}
    {\it Top:} 
    The time of each Kepler transit was measured, along with the local
    slope of the light curve.  The two quantities appear
    anti-correlated, which is most easily explained by starspot
    crossings during the first (second) half of transit inducing a
    positive (negative) TTV, provided that the orbit is prograde
    \citep{mazeh_time_2015}.  The units along the abscissa are most
    easily understood by considering that the stellar flux changes by
    $\sim$60\,ppt per half rotation period
    ($\sim$1.3\,days).
    {\it Bottom:} Radial velocities acquired using Keck/HIRES during
    the transit of 2021 Aug 7, observed simultaneously with MuSCAT3.
    The best-fit sky-projected obliquity for this model is listed in
    red.  While the intra-exposure RV scatter precludes a detection of
    the RM effect, the photometric transit depths are consistent
    across the {\it griz} bandpasses.
    \label{fig:obliquity}
	}
\end{figure}
\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.8\textwidth]{f14.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Radii, orbital periods, and ages of transiting exoplanets}.
    Planets older than $10^9$ years are shown with gray markers.
    Planets younger than a gigayear with ${\rm \tau}/\sigma_{\rm \tau}
    > 3$ are emphasized, for $\tau$ the age.  \pn\ is shown with a
    star.  The large sizes of the youngest transiting planets could be
    explained by their primordial atmospheres not yet having
    evaporated; direct measurements of the atmospheric outflows or
    planetary masses would help to confirm this expectation.
    Selection effects may also be important.  Parameters are from the
    NASA Exoplanet Archive (2021 Sept 15).
		\label{fig:rp_period_age}
	}
\end{figure*}

@@

If the \pn\ transit signal is created by a genuine planet,  then to
our knowledge its age would make it the youngest planet yet found by
the main Kepler mission.\footnote{The re-purposed K2 mission however
has found two younger systems containing five planets: K2-33b
($9\pm1\,{\rm Myr}$; \citealt{Mann_K2_33b_2016,David_et_al_2017}) and
V1298 Tau ($23\pm4\,{\rm Myr}$; \citealt{david_four_2019}).}  A
portion of the light curve is shown with the phase-folded transit in
Figure~\ref{fig:lc}.  The \texttt{PDCSAP} photometry is dominated by a
quasi-periodic starspot signal with a peak-to-peak amplitude that
varies between 2\% and 8\%, depending on the degree of asymmetry in
spot coverage between the stellar hemispheres.  The starspot signal
repeats every $2.642\pm0.042$\,days, where the uncertainty is based on
the scatter observed between different Kepler quarters.  The details
of how we model the starspot signal to derive the parameters of the
planet are discussed in Appendix~\ref{app:gptransit}.  The stellar
parameters are listed in Table~\ref{tab:starparams}; the stellar mass,
radius, and effective temperature are found by interpolating against a
38$\,$Myr MIST isochrone.  The systematic uncertainties are taken to
be the difference in parameters from the PARSEC isochrones, and the
reported uncertainties are a quadrature sum of the statistical and
systematic components.  The transit-fitting results, given in
Table~\ref{tab:posterior}, are consistent with a mini-Neptune sized
planet ($3.34\pm 0.18\,R_\oplus$) on a close-in circular\footnote{Our
transit fitting yields $e<0.48$ at 2-$\sigma$; the constraints on the
eccentricity are not particularly strong.} orbit around a G8V star
($0.88 \pm 0.02 R_\odot$).

Could the transit signal be produced by anything other than a planet
orbiting this near-solar analog?  \citet{morton_false_2016} validated
the planet based on the transit shape, arguing that the most probable
false positive scenario was that of a background eclipsing binary,
which had a model-dependent probability of $\approx10^{-5}$.  As we
describe in Appendix~\ref{app:companionstar} however, Kepler\,1627A is
accompanied by a low-mass stellar companion ($M_{\rm B}\approx
0.33\,M_\odot$).  We therefore reconsidered this false positive
scenario in some detail.

@@

The expectation however is for the spot-induced signals to be smooth,
at worst with contributions at $0.5\,P_{\rm rot}$ or $0.25\,P_{\rm
rot}$ \citep{klein_simulated_2020}.  We therefore fitted the RVs using
the \citet{hirano_analytic_2010,hirano_2011} models for the
Rossiter-McLaughlin (RM) effect, and allowed for an optional linear
and quadratic trend in time to fit the spot-induced signal.  We
followed the methodology developed by \citet{stefansson_2020}.  We
allowed the sky-projected obliquity, the projected stellar equatorial
velocity, and the Gaussian dispersion of the spectral lines to vary,
and fixed the limb-darkening using the $V$-band tabulation from
\citet{claret_gravity_2011}.  We assumed a prior on $v\sin i$ and
$a/R_\star$ from Table~\ref{tab:starparams}, and also allowed for a
white-noise jitter term to be added in quadrature to the measurement
uncertainties.

@@

The presence of this additional scatter prevents a clear detection of
the RM effect.  The reason can be understood through a
model-comparison analysis.  If we compare the model with a quadratic trend and
the RM effect (Figure~\ref{fig:obliquity}) against a model with a
linear trend and the RM effect, or even a model with no RM effect at
all, then the respective Bayesian Information Criterion (BIC) values
are as follows.
% from logs/rm_freejitter_FINAL.log
\begin{align}
  {\rm BIC} &= 251.7  \quad ({\rm Quadratic+RM}) \nonumber \\
  {\rm BIC} &= 257.7  \quad ({\rm Linear+RM}) \nonumber \\
  {\rm BIC} &= 247.1  \quad ({\rm Only\ Quadratic}).
\end{align}
There is therefore no evidence to prefer the model with the RM effect
against a model that only accounts for the stellar variability.
The ``only quadratic'' model does particularly well because it can
inflate the jitter term to account for any additional scatter during
the transit, and it has fewer free parameters.  However, we cannot
justify a physical prior on the jitter term, because we do not
understand the origin of the exposure-to-exposure scatter.  As noted
above, the velocity deviations from starspots are expected to have
contributions at the stellar rotation frequency, or harmonics thereof.
This jitter is present on the exposure timescale (15 minutes), which
is only 0.4\% of the stellar rotation period; it is not obvious that
starspots could be the culprit.

The amplitude of both the spot-induced trend and the jitter are larger
than recent comparable measurements in systems such as AU~Mic
\citep{palle_transmission_2020}, DS~Tuc
\citep{montet_young_2020,zhou_well_2020} and TOI~942
\citep{wirth_2021_toi942}.  One possible explanation for the jitter is
that it is astrophysical in origin, and that it is caused by some
novel process operating on the surface of Kepler 1627A.  Another
possibility is that our RV analysis underestimates the measurement
uncertainties; the S/N per resolution element for each of our spectra
(70 to 80) was moderated by the observing requirements, and the rapid
rotation of the star could affect the velocity extraction.
Observations at higher precision are necessary to distinguish these
two possibilities, and remain worthwhile as a means toward clarifying
the orbital geometry of \pn.

>>>>

\keywords{
  exoplanet evolution (491),
  %stellar associations (1582),
  open star clusters (1160),
	stellar ages (1581)
}

@@

The K2 and TESS missions have now enabled the detection of about ten
close-in planets younger than 100 million years, all smaller than
Jupiter
\citep{Mann_K2_33b_2016,David_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,plavchan_planet_2020,rizzuto_tess_2020,martioli_aumicbc_2021}.
The Kepler mission however has not yielded any planets with precise
ages below one gigayear \citep{Meibom_et_al_2013}.  The reason is that
during the main Kepler mission (2009--2013), only four open clusters
were known in the Kepler field,
%: NGC\,6866, NGC\,6811, NGC\,6819, and NGC\,6791,
with ages spanning 0.7\,Gyr to 9\,Gyr
\citep{meibom_kepler_2011}.  Though isochronal, gyrochronal, and lithium-based
analyses suggest that younger Kepler planets do exist
\citep{berger_identifying_2018,david_sizes_2021}, accurate and precise
age measurements typically require an ensemble of stars.  Fortunately,
recent analyses of the kinematic, photometric, and astrometric Gaia
data have since expanded our knowledge of cluster memberships
\citep[{\it
e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021,Kerr2021}.
As part of our Cluster Difference Imaging Photometric Survey (CDIPS,
\citealt{bouma_cdipsI_2019}), we concatenated the available analyses
from the literature, which yielded a list of candidate young and
age-dated stars (see Appendix~\ref{app:targetlist}).

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.5\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.6cm}
	\caption{
		{\bf Kepler and TESS views of the $\delta$~Lyr cluster.} Colored
		points are kinematically selected candidate $\delta$\,Lyr cluster
		members.  Both Kepler (gray panels) and TESS observed portions of
		the cluster.  Naked-eye stars ($m_{\rm V}<6.5$) are shown in gray;
		Deneb, Vega, and $\delta$~Lyr are orange.  Kepler\,1627 (green star) was observed during the
		entirety of the Kepler mission.  Colors of candidate cluster
		members indicate the duration of TESS observations to date, in
		units of lunar months.
		\label{fig:skychart}
	}
\end{figure}

@@

@@

		\subfloat{
			\includegraphics[width=0.5\textwidth]{f5a.pdf}
			\includegraphics[width=0.5\textwidth]{f5b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The light curve of Kepler\,1627.}
    {\it Top}: 
    The Kepler data span 1{,}437 days (3.9 years), sampled at
    30 minute cadence;  a 100 day segment is shown.  The
    top panel shows the \texttt{PDCSAP} median-subtracted flux in
    units of parts-per-thousand ($\times 10^{-3}$).  The dominant
    signal is induced by starspots.  The stellar
    variability model (orange line) is subtracted below, revealing the
    transits of \pn, as well as other deviations from the model.  The
    online Figure Set spans the entire 3.9 years of observations.
    {\it Bottom}:
    Phase-folded transit of Kepler 1627Ab with stellar variability
    removed.  Windows over 20 hours ({\it left}) and the entire orbit
    ({\it right}) are shown.  The 2-$\sigma$ uncertainties and the
    maximum best-fit model are the faint purple band, and the dark
    purple line.  Gray points are individual flux measurements; black
    points bin these every 20 minutes, and have a representative error bar
    shown.  The asymmetric residual during transit appears to be robust when using
    different
    detrending methods.
    \label{fig:lc}
  }
\end{figure*}

@@

		\subfloat{
			\includegraphics[width=0.49\textwidth]{f16.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Evidence for a prograde orbit of Kepler 1627 Ab.}
    {\it Top:} 
    The time of each Kepler transit was measured, along with the local
    slope of the light curve.  The two quantities appear
    anti-correlated, which is most easily explained by starspot
    crossings during the first (second) half of transit inducing a
    positive (negative) TTV, provided that the orbit is prograde
    \citep{mazeh_time_2015}.  The units along the abscissa are most
    easily understood by considering that the stellar flux changes by
    $\sim$60\,ppt per half rotation period
    ($\sim$1.3\,days).
    {\it Bottom:} Radial velocities acquired using Keck/HIRES during
    the transit of 2021 Aug 7, observed simultaneously with MuSCAT3.
    The best-fit sky-projected obliquity for this model is listed in
    red.  While the intra-exposure RV scatter precludes a detection of
    the RM effect, the photometric transit depths are consistent
    across the {\it griz} bandpasses.
    \label{fig:obliquity}
	}
\end{figure}
\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.8\textwidth]{f14.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Radii, orbital periods, and ages of transiting exoplanets}.
    %Planets older than $10^9$ years are shown with gray markers.
    Planets younger than a gigayear with ${\rm \tau}/\sigma_{\rm \tau}
    > 3$ are emphasized, for $\tau$ the age.  \pn\ is shown with a
    star.  The large sizes of the youngest transiting planets could be
    explained by their primordial atmospheres not yet having
    evaporated; direct measurements of the atmospheric outflows or
    planetary masses would help to confirm this expectation.
    Selection effects may also be important.  Parameters are from the
    NASA Exoplanet Archive (2021 Sept 15).
		\label{fig:rp_period_age}
	}
\end{figure*}

@@

If the \pn\ transit signal is created by a genuine planet,  then to
our knowledge its age would make it the youngest planet yet found by
the main Kepler mission.\footnote{The re-purposed K2 mission however
has found two younger systems containing five planets: K2-33b
($9\pm1\,{\rm Myr}$; \citealt{Mann_K2_33b_2016,David_et_al_2017}) and
V1298 Tau ($23\pm4\,{\rm Myr}$; \citealt{david_four_2019}).}  The light
curve and the phase-folded transit are shown in
Figure~\ref{fig:lc}.  The photometry is dominated by a
quasi-periodic starspot signal with a peak-to-peak amplitude that
varies between 2\% and 8\%.  The signal
repeats every $2.642\pm0.042$\,days, where the quoted uncertainty is based on
the scatter in rotation periods measured from each individual Kepler quarter. 
Our model for the starspot signal is discussed in detail
in Appendix~\ref{app:gptransit}, and the adopted stellar
parameters are listed in Table~\ref{tab:starparams}.  The stellar mass,
radius, and effective temperature are found by interpolating against a
38$\,$Myr MIST isochrone.  The systematic uncertainties are taken
to be the difference between the PARSEC and MIST isochrones, and the
reported uncertainties are a quadrature sum of the statistical and
systematic components. 
The transit fitting results, given in
Table~\ref{tab:posterior}, are consistent with a mini-Neptune sized
planet ($3.34\pm 0.18\,R_\oplus$) on a close-in circular\footnote{Our
transit fitting yields $e<0.48$ at 2-$\sigma$; the constraints on the
eccentricity are not particularly strong.} orbit around a G8V star
($0.88 \pm 0.02 R_\odot$).

Could this transit signal be produced by anything other than a planet
orbiting this near-solar analog?  \citet{morton_false_2016} validated
the planet based on the transit shape, arguing that the most probable
false positive scenario was that of a background eclipsing binary,
which had a model-dependent probability of $\approx10^{-5}$.  As we
describe in Appendix~\ref{app:companionstar} however, Kepler\,1627A is
accompanied by a low-mass stellar companion ($M_{\rm B}\approx
0.33\,M_\odot$).  We therefore reassessed false positive
scenarios in some detail.

@@

The expectation however is for the starspot-induced signals to be smooth,
at worst with contributions at $0.5\,P_{\rm rot}$ or $0.25\,P_{\rm
rot}$ \citep{klein_simulated_2020}.  We therefore fitted the RVs using
the \citet{hirano_analytic_2010,hirano_2011} models for the
Rossiter-McLaughlin (RM) effect, and allowed for an optional linear
and quadratic trend in time to fit the 800\ms\ spot-induced trend.  We
followed the methodology developed by \citet{stefansson_2020}.  We
allowed the sky-projected obliquity, the projected stellar equatorial
velocity, and the Gaussian dispersion of the spectral lines to vary,
and fixed the limb-darkening using the $V$-band tabulation from
\citet{claret_gravity_2011}.  We assumed a prior on $v\sin i$ and
$a/R_\star$ from Table~\ref{tab:starparams}, and also allowed for a
white-noise jitter term to be added in quadrature to the measurement
uncertainties.

@@

The presence of this additional scatter prevents a clear detection of
the RM effect.  The reason can be understood via
model comparison.  If we compare the model with a quadratic trend and
the RM effect (Figure~\ref{fig:obliquity}) against a model with a
linear trend and the RM effect, or even a model with no RM effect at
all, then the respective Bayesian Information Criterion (BIC) values
are as follows.
% from logs/rm_freejitter_FINAL.log
\begin{align}
  {\rm BIC} &= 251.7  \quad ({\rm Quadratic+RM}) \nonumber \\
  {\rm BIC} &= 257.7  \quad ({\rm Linear+RM}) \nonumber \\
  {\rm BIC} &= 247.1  \quad ({\rm Only\ Quadratic}).
\end{align}
There is therefore no evidence to prefer the model with the RM effect
against a model that only accounts for the stellar variability.
The ``only quadratic'' model does particularly well because it can
inflate the jitter term to account for scatter during
the transit (even if the scatter contains astrophysics!), and it has fewer free parameters.  However, we cannot
justify a physical prior on the jitter term, because we do not
understand the origin of the exposure-to-exposure scatter.  As noted
above, the velocity deviations from starspots are expected to have
contributions at the stellar rotation frequency, or harmonics thereof.
This jitter is present on the exposure timescale (15 minutes), which
is only 0.4\% of the stellar rotation period; it is not obvious that
starspots would be the culprit.

The amplitude of both the spot-induced trend and the jitter are larger
than recent comparable measurements in systems such as AU~Mic
\citep{palle_transmission_2020}, DS~Tuc
\citep{montet_young_2020,zhou_well_2020} and TOI~942
\citep{wirth_2021_toi942}.  One possible explanation for the jitter is
that it is astrophysical in origin, and that it is caused by some
novel process operating on the surface of Kepler 1627A.  Another
possibility is that our RV analysis underestimates the measurement
uncertainties; the S/N per resolution element for each of our spectra
was moderated by the observing requirements to 70 to 80, and the rapid
rotation of the star could affect accuracy of the uncertainties from
the velocity extraction.  Observations at higher precision are
necessary to distinguish these two possibilities, and remain
worthwhile as a means toward clarifying the orbital geometry of \pn.
